### 2016-07-17 P4-Explore-and-Summarize-Data/submission.Rmd
Update submission.Rmd

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
rw <- read.csv('wineQualityReds.csv')
```

# Statistical Summary 
### Structure of the dataset
```{r echo=FALSE, structure}
str(rw)
```

@@

### Statistical summary of the variables in the dataset
```{r echo=FALSE, summary}
summary(rw)
```
As we are primarily interested in quality of the red wines, let's see some basic statistics about it
```{r echo=FALSE}
summary(rw$quality)
```
* We've found that the quality lies in the range of [3, 8] with mean is equal to 5.636.

# Univariate Plots Section
```{r echo=FALSE, Univariate_Plots, message=FALSE}
p1 <- qplot(data = rw, fixed.acidity)
p2 <- qplot(data = rw, volatile.acidity)
p3 <- qplot(data = rw, citric.acid)
p4 <- qplot(data = rw, residual.sugar)
p5 <- qplot(data = rw, chlorides)
p6 <- qplot(data = rw, free.sulfur.dioxide)
p7 <- qplot(data = rw, total.sulfur.dioxide)
p8 <- qplot(data = rw, density)
p9 <- qplot(data = rw, pH)
p10 <- qplot(data = rw, sulphates)
p11 <- qplot(data = rw, alcohol)
p12 <- qplot(data = rw, quality)

@@

```{r echo=FALSE}
rw$rating <- ifelse(rw$quality < 5, 'poor', ifelse(
  rw$quality < 7, 'good', 'ideal'))

@@

```{r echo=FALSE, message=FALSE}
rw$total.acidity <- rw$fixed.acidity + rw$volatile.acidity + rw$citric.acid

@@

```{r echo=FALSE, message=FALSE}
ggplot(data = rw, aes(x = fixed.acidity)) +
  geom_histogram() + 
  scale_x_log10()

ggplot(data = rw, aes(x = volatile.acidity)) +
  geom_histogram() + 
  scale_x_log10()

ggplot(data = rw, aes(x = free.sulfur.dioxide)) +
  geom_histogram() + 
  scale_x_log10()

ggplot(data = rw, aes(x = total.sulfur.dioxide)) +
  geom_histogram() + 
  scale_x_log10()

ggplot(data = rw, aes(x = sulphates)) +
  geom_histogram() + 
  scale_x_log10()

ggplot(data = rw, aes(x = alcohol)) +
  geom_histogram() + 
  scale_x_log10()
```

@@

```{r echo=FALSE, message=FALSE}
length(subset(rw, citric.acid == 0)$citric.acid)
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = rw, aes(x = residual.sugar)) +
  geom_histogram() +
  scale_x_continuous(lim = c(0.5, quantile(rw$residual.sugar, 0.95)))

ggplot(data = rw, aes(x = chlorides)) +
  geom_histogram() +
  scale_x_continuous(lim = c(0.04, quantile(rw$chlorides, 0.95)))
```

@@

```{r echo=FALSE, message=FALSE}
str(rw)
```

@@

# Bivariate Plots Section
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
bivariate_plot <- function(y, ylab) {
  return (qplot(data = rw, x = '', y = y, 
                geom = 'boxplot', 
                xlab = '', 
                ylab = ylab))
}

grid.arrange(bivariate_plot(rw$fixed.acidity, 'fixed.acidity'),
             bivariate_plot(rw$volatile.acidity, 'volatile.acidity'),
             bivariate_plot(rw$citric.acid, 'citric.acid'),
             bivariate_plot(rw$residual.sugar, 'residual.sugar'),
             bivariate_plot(rw$chlorides, 'chlorides'),
             bivariate_plot(rw$free.sulfur.dioxide, 'free.sulfur.dioxide'),
             bivariate_plot(rw$total.sulfur.dioxide, 'total.sulfur.dioxide'),
             bivariate_plot(rw$density, 'density'),
             bivariate_plot(rw$pH, 'pH'),
             bivariate_plot(rw$sulphates, 'sulphates'),
             bivariate_plot(rw$alcohol, 'alcohol'),
             ncol = 4)
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
quality_plot <- function(x, y, ylab) {
  return (ggplot(data = rw, aes(x, y)) +
           geom_boxplot(fill = 'green') +
            xlab('quality') +
            ylab(ylab))
}

grid.arrange(quality_plot(rw$quality, rw$fixed.acidity, 'fixed.acidity'),
             quality_plot(rw$quality, rw$volatile.acidity, 'volatile.acidity'),
             quality_plot(rw$quality, rw$citric.acid, 'citric.acid'),
             quality_plot(rw$quality, rw$residual.sugar, 'residual.sugar'),
             quality_plot(rw$quality, rw$chlorides, 'chlorides'),
             quality_plot(rw$quality, rw$free.sulfur.dioxide, 'free.sulfur.dioxide'),
             quality_plot(rw$quality, rw$total.sulfur.dioxide, 'total.sulfur.dioxide'),
             quality_plot(rw$quality, rw$density, 'density'),
             quality_plot(rw$quality, rw$pH, 'pH'),
             quality_plot(rw$quality, rw$sulphates, 'sulphates'),
             quality_plot(rw$quality, rw$alcohol, 'alcohol'),
             ncol = 4)
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
rating_plot <- function(x, y, ylab) {
  return (ggplot(data = rw, aes(x, y)) +
           geom_boxplot(fill = 'orange') +
            xlab('rating') +
            ylab(ylab))
}

grid.arrange(rating_plot(rw$rating, rw$fixed.acidity, 'fixed.acidity'),
             rating_plot(rw$rating, rw$volatile.acidity, 'volatile.acidity'),
             rating_plot(rw$rating, rw$citric.acid, 'citric.acid'),
             rating_plot(rw$rating, rw$residual.sugar, 'residual.sugar'),
             rating_plot(rw$rating, rw$chlorides, 'chlorides'),
             rating_plot(rw$rating, rw$free.sulfur.dioxide, 'free.sulfur.dioxide'),
             rating_plot(rw$rating, rw$total.sulfur.dioxide, 'total.sulfur.dioxide'),
             rating_plot(rw$rating, rw$density, 'density'),
             rating_plot(rw$rating, rw$pH, 'pH'),
             rating_plot(rw$rating, rw$sulphates, 'sulphates'),
             rating_plot(rw$rating, rw$alcohol, 'alcohol'),
             ncol = 4)
```

@@

* low ```fixed.acidity``` and ```volatile.acidity```
* higher ```sulfur.dioxide``` and ```sulphates```
* low ```pH``` and higher ```density```

@@

* higher ```fixed.acidity``` , lower ```volatile.acidity``` and higher ```citric.acid```
* low ```sulfur.dioxide``` and ```sulphates```
* lower ```pH``` and ```density```

@@

```{r echo=FALSE, message=FALSE}
correlations <- c(cor.test(rw$fixed.acidity, rw$quality)$estimate,
                  cor.test(rw$volatile.acidity, rw$quality)$estimate,
                  cor.test(rw$citric.acid, rw$quality)$estimate,
                  cor.test(rw$residual.sugar, rw$quality)$estimate,
                  cor.test(rw$chlorides, rw$quality)$estimate,
                  cor.test(rw$free.sulfur.dioxide, rw$quality)$estimate,
                  cor.test(rw$total.sulfur.dioxide, rw$quality)$estimate,
                  cor.test(rw$density, rw$quality)$estimate,
                  cor.test(rw$pH, rw$quality)$estimate,
                  cor.test(rw$sulphates, rw$quality)$estimate,
                  cor.test(rw$alcohol, rw$quality)$estimate,
                  cor.test(rw$total.acidity, rw$quality)$estimate)

@@

```{r echo=FALSE, message=FALSE}
ggplot(data = rw, aes(log10(total.acidity), density)) +
  geom_point() +
  scale_x_continuous() +
  geom_smooth()

ggplot(data = rw, aes(log10(total.acidity), pH)) +
  geom_point() +
  scale_x_continuous() +
  geom_smooth()

@@

```{r echo=FALSE, message=FALSE, Multivariate_Plots}
ggplot(data = rw, aes(pH, alcohol, color = quality)) +
  geom_point() +
  facet_wrap(~rating)

ggplot(data = rw, aes(alcohol, citric.acid, color = quality)) +
  geom_point() +
  facet_wrap(~rating)

ggplot(data = rw, aes(alcohol, sulphates, color = quality)) +
  geom_point() +
  facet_wrap(~rating)

ggplot(data = rw, aes(citric.acid, fixed.acidity, color = quality)) +
  geom_point() +
  facet_wrap(~rating)
```

@@

### Plot One : Effect of alcohol on wine quality
```{r echo=FALSE, message=FALSE, Plot_One}
ggplot(data = rw, aes(quality, alcohol, fill = rating)) +
  geom_boxplot() +
  ggtitle('Alcohol Levels on Wine Quality') +
  xlab('Quality') +
  ylab('Alcohol (% volume)')

@@

### Description One
Above plot clearly shows higher % of alcohol yields better wine.

### Plot Two : Effect of acids on wine quality
```{r echo=FALSE, message=FALSE, Plot_Two}
grid.arrange(ggplot(data = rw, aes(quality, fixed.acidity, fill = rating)) +
               geom_boxplot(),
             ggplot(data = rw, aes(quality, volatile.acidity, fill = rating)) +
               geom_boxplot(),
             ggplot(data = rw, aes(quality, citric.acid, fill = rating)) +
               geom_boxplot(),
             ncol = 2)
```

### Description Two
```fixed.acidity``` and ```citric.acid``` have highly correlated with quality but ```volatile.acidity``` has negative impact on quality.

### Plot Three : Effect of pH on wine quality
```{r echo=FALSE, message=FALSE, Plot_Three}
ggplot(data = rw, aes(quality, pH, fill = rating)) +
  geom_boxplot() +
  ggtitle('pH Levels on Wine Quality')

@@

### Description Three
Lower the pH, better the wine. This also substantiate that higher the acidic tends to better the wine.

@@

```{r echo=FALSE, message=FALSE} 
ggplot(data = subset(rw, rating != 'good'), aes(alcohol, sulphates, color = rating)) +
  geom_point() +
  ggtitle('Alcohol vs Sulphates on Wine Quality') +
  xlab('Sulphates (g / dm^3)') +
  ylab('Alcohol (% volume)')
```

@@

>>>>

```{r echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r }
library(ggplot2)
library(dplyr)
library(gridExtra)
```

```{r, Load_the_Data}
rw <- read.csv('wineQualityReds.csv')
```

# Statistical Summary 
### Structure of the dataset
```{r , structure}
str(rw)
```

@@

### Statistical summary of the variables in the dataset
```{r, summary}
summary(rw)
```
As we are primarily interested in quality of the red wines, let's see some basic statistics about it
```{r }
summary(rw$quality)
```
* We've found that the quality lies in the range of [3, 8] with mean is equal to 5.636.

# Univariate Plots Section
```{r, Univariate_Plots}
p1 <- qplot(data = rw, fixed.acidity, xlab = 'fixed.acidity (g / dm^3)')
p2 <- qplot(data = rw, volatile.acidity, xlab = 'volatile.acidity (g / dm^3)')
p3 <- qplot(data = rw, citric.acid, xlab = 'citric.acid (g / dm^3)')
p4 <- qplot(data = rw, residual.sugar, xlab = 'residual.sugar (g / dm^3)')
p5 <- qplot(data = rw, chlorides, xlab = 'chlorides (g / dm^3)')
p6 <- qplot(data = rw, free.sulfur.dioxide, xlab = 'free.sulfur.dioxide (mg / dm^3)')
p7 <- qplot(data = rw, total.sulfur.dioxide, xlab = 'total.sulfur.dioxide (mg / dm^3)')
p8 <- qplot(data = rw, density, xlab = 'density (g / dm^3)')
p9 <- qplot(data = rw, pH)
p10 <- qplot(data = rw, sulphates, xlab = 'sulpahtes (g / dm^3)')
p11 <- qplot(data = rw, alcohol, xlab = 'alcohol (% volume)')
p12 <- qplot(data = rw, quality)

@@

```{r }
rw$rating <- ifelse(rw$quality < 5, 'poor', ifelse(
  rw$quality < 7, 'good', 'ideal'))

@@

```{r }
rw$total.acidity <- rw$fixed.acidity + rw$volatile.acidity + rw$citric.acid

@@

```{r }
ggplot(data = rw, aes(x = fixed.acidity)) +
  geom_histogram() + 
  scale_x_log10() +
  xlab('fixed.acidity (g / dm^3)')

ggplot(data = rw, aes(x = volatile.acidity)) +
  geom_histogram() + 
  scale_x_log10() +
  xlab('volatile.acidity (g / dm^3)')

ggplot(data = rw, aes(x = free.sulfur.dioxide)) +
  geom_histogram() + 
  scale_x_log10() +
  xlab('free.sulfur.dioxide (mg / dm^3)')

ggplot(data = rw, aes(x = total.sulfur.dioxide)) +
  geom_histogram() + 
  scale_x_log10() +
  xlab('total.sulfur.dioxide (mg / dm^3)')

ggplot(data = rw, aes(x = sulphates)) +
  geom_histogram() + 
  scale_x_log10() +
  xlab('sulphates (g / dm^3)')

ggplot(data = rw, aes(x = alcohol)) +
  geom_histogram() + 
  scale_x_log10() +
  xlab('alcohol (% volume')
```

@@

```{r }
length(subset(rw, citric.acid == 0)$citric.acid)
```

@@

```{r }
ggplot(data = rw, aes(x = residual.sugar)) +
  geom_histogram() +
  scale_x_continuous(lim = c(0.5, quantile(rw$residual.sugar, 0.95))) +
  xlab('residual.sugar (g / dm^3)')

ggplot(data = rw, aes(x = chlorides)) +
  geom_histogram() +
  scale_x_continuous(lim = c(0.04, quantile(rw$chlorides, 0.95))) +
  xlab('chlorides (g / dm^3)')
```

@@

```{r }
str(rw)
```

@@

# Bivariate Plots Section
```{r , Bivariate_Plots}
bivariate_plot <- function(y, ylab) {
  return (qplot(data = rw, x = '', y = y, 
                geom = 'boxplot', 
                xlab = '', 
                ylab = ylab))
}

grid.arrange(bivariate_plot(rw$fixed.acidity, 'fixed.acidity (g / dm^3)'),
             bivariate_plot(rw$volatile.acidity, 'volatile.acidity (g / dm^3)'),
             bivariate_plot(rw$citric.acid, 'citric.acid (g / dm^3)'),
             bivariate_plot(rw$residual.sugar, 'residual.sugar (g / dm^3)'),
             bivariate_plot(rw$chlorides, 'chlorides (g / dm^3)'),
             bivariate_plot(rw$free.sulfur.dioxide, 'free.sulfur.dioxide (mg / dm^3)'),
             bivariate_plot(rw$total.sulfur.dioxide, 'total.sulfur.dioxide (mg / dm^3)'),
             bivariate_plot(rw$density, 'density (g / cm^3)'),
             bivariate_plot(rw$pH, 'pH'),
             bivariate_plot(rw$sulphates, 'sulphates (g / dm^3)'),
             bivariate_plot(rw$alcohol, 'alcohol (% volume)'),
             ncol = 4)
```

@@

```{r }
quality_plot <- function(x, y, ylab) {
  return (ggplot(data = rw, aes(x, y)) +
           geom_boxplot(fill = 'green') +
            xlab('quality') +
            ylab(ylab))
}

grid.arrange(quality_plot(rw$quality, rw$fixed.acidity, 'fixed.acidity (g / dm^3)'),
             quality_plot(rw$quality, rw$volatile.acidity, 
                          'volatile.acidity (g / dm^3)'),
             quality_plot(rw$quality, rw$citric.acid, 'citric.acid (g / dm^3)'),
             quality_plot(rw$quality, rw$residual.sugar, 'residual.sugar (g / dm^3)'),
             quality_plot(rw$quality, rw$chlorides, 'chlorides (g / dm^3)'),
             quality_plot(rw$quality, rw$free.sulfur.dioxide, 
                          'free.sulfur.dioxide (mg / dm^3)'),
             quality_plot(rw$quality, rw$total.sulfur.dioxide, 
                          'total.sulfur.dioxide (mg / dm^3)'),
             quality_plot(rw$quality, rw$density, 
                          'density (g / cm^3)'),
             quality_plot(rw$quality, rw$pH, 'pH'),
             quality_plot(rw$quality, rw$sulphates, 'sulphates (g / dm^3)'),
             quality_plot(rw$quality, rw$alcohol, 'alcohol (% volume)'),
             ncol = 4)
```

@@

```{r }
rating_plot <- function(x, y, ylab) {
  return (ggplot(data = rw, aes(x, y)) +
           geom_boxplot(fill = 'orange') +
            xlab('rating') +
            ylab(ylab))
}

grid.arrange(rating_plot(rw$rating, rw$fixed.acidity, 'fixed.acidity (g / dm^3)'),
             rating_plot(rw$rating, rw$volatile.acidity, 
                         'volatile.acidity (g / dm^3)'),
             rating_plot(rw$rating, rw$citric.acid, 'citric.acid (g / dm^3)'),
             rating_plot(rw$rating, rw$residual.sugar, 'residual.sugar (g / dm^3)'),
             rating_plot(rw$rating, rw$chlorides, 'chlorides (g / dm^3)'),
             rating_plot(rw$rating, rw$free.sulfur.dioxide, 
                         'free.sulfur.dioxide (mg / dm^3)'),
             rating_plot(rw$rating, rw$total.sulfur.dioxide, 
                         'total.sulfur.dioxide (mg / dm^3)'),
             rating_plot(rw$rating, rw$density, 'density (g / cm^3)'),
             rating_plot(rw$rating, rw$pH, 'pH'),
             rating_plot(rw$rating, rw$sulphates, 'sulphates (g / dm^3)'),
             rating_plot(rw$rating, rw$alcohol, 'alcohol (% volume)'),
             ncol = 4)
```

@@

* low ```fixed.acidity``` and ```volatile.acidity```
* higher ```sulfur.dioxide```
* low ```pH``` and higher ```density```

@@

* higher ```fixed.acidity``` , lower ```volatile.acidity``` and higher ```citric.acid```
* low ```sulfur.dioxide``` and higher ```sulphates```
* lower ```pH``` and ```density```

@@

```{r }
correlations <- c(cor.test(rw$fixed.acidity, rw$quality)$estimate,
                  cor.test(rw$volatile.acidity, rw$quality)$estimate,
                  cor.test(rw$citric.acid, rw$quality)$estimate,
                  cor.test(rw$residual.sugar, rw$quality)$estimate,
                  cor.test(rw$chlorides, rw$quality)$estimate,
                  cor.test(rw$free.sulfur.dioxide, rw$quality)$estimate,
                  cor.test(rw$total.sulfur.dioxide, rw$quality)$estimate,
                  cor.test(rw$density, rw$quality)$estimate,
                  cor.test(rw$pH, rw$quality)$estimate,
                  cor.test(rw$sulphates, rw$quality)$estimate,
                  cor.test(rw$alcohol, rw$quality)$estimate,
                  cor.test(rw$total.acidity, rw$quality)$estimate)

@@

```{r }
ggplot(data = rw, aes(log10(total.acidity), density)) +
  geom_point(alpha = 1/10) +
  scale_x_continuous() +
  geom_smooth() +
  xlab('log10[total.acidity (g / dm^3)]') +
  ylab('density (g / dm^3)')

ggplot(data = rw, aes(log10(total.acidity), pH)) +
  geom_point(alpha = 1/10) +
  scale_x_continuous() +
  geom_smooth() +
  xlab('log10[total.acidity (g / dm^3)]')

@@

```{r , Multivariate_Plots}
ggplot(data = rw, aes(pH, alcohol, color = quality)) +
  geom_point() +
  facet_wrap(~rating) +
  ylab('alcohol (% volume)')

ggplot(data = rw, aes(alcohol, citric.acid, color = quality)) +
  geom_point() +
  facet_wrap(~rating) +
  xlab('alcohol (% volume)') +
  ylab('citric.acid (g / dm^3)')

ggplot(data = rw, aes(alcohol, sulphates, color = quality)) +
  geom_point() +
  facet_wrap(~rating) +
  xlab('alcohol (% volume)') +
  ylab('sulphates (g / dm^3)')

ggplot(data = rw, aes(citric.acid, fixed.acidity, color = quality)) +
  geom_point() +
  facet_wrap(~rating) +
  xlab('citric.acid (g / dm^3)') +
  ylab('fixed.acid (g / dm^3)')
```

@@

### Plot One : Effect of alcohol on wine quality
```{r , Plot_One}
ggplot(data = rw, aes(quality, alcohol, fill = rating)) +
  geom_boxplot() +
  ggtitle('Alcohol Levels on Wine Quality') +
  xlab('Quality') +
  ylab('Alcohol (% volume)')

@@

### Description One
As alcohol is highly correlated with the quality, it is better to see its pattern with varying rating. From the above plot, it clearly shows higher % of alcohol yields better wine.

### Plot Two : Effect of acids on wine quality
```{r , Plot_Two}
grid.arrange(ggplot(data = rw, aes(quality, fixed.acidity, fill = rating)) +
               geom_boxplot(),
             ggplot(data = rw, aes(quality, volatile.acidity, fill = rating)) +
               geom_boxplot(),
             ggplot(data = rw, aes(quality, citric.acid, fill = rating)) +
               geom_boxplot(),
             ncol = 2)
```

### Description Two
As more the acidic better is the wine. It would be better to see which acids have more impact on wine quality. Above plot shows, ```fixed.acidity``` and ```citric.acid``` have highly correlated with quality but ```volatile.acidity``` has negative impact on quality.

### Plot Three : Effect of pH on wine quality
```{r , Plot_Three}
ggplot(data = rw, aes(log10(total.acidity), pH)) +
  geom_point(alpha = 1/10) +
  scale_x_continuous() +
  geom_smooth() +
  xlab('total.acidity (g / dm^3)') +
  ggtitle('pH Relation on Wine Quality')

@@

### Description Three
From the above plot, it shows there is linear relationships between pH and total.acidity. This means lower the pH, more acidic the wines. This also substantiate that higher the acidic tends to better the wine.

@@

```{r } 
ggplot(data = subset(rw, rating != 'good'), aes(alcohol, sulphates, color = rating)) +
  geom_point() +
  ggtitle('Alcohol vs Sulphates on Wine Quality') +
  xlab('Sulphates (g / dm^3)') +
  ylab('Alcohol (% volume)')
```

### Description Four
It would be great to see the real pattern between good and bad wines. Above plot differentiate between good and bad wines. It shows higher the % of alcohol and higher the sulphates give better wines.

@@

One thing that is still unclear is the amount of ```residual.sugar```. It contains many outliers, also after doing some operation we get its common range from 1 to 4. But we can't find its amount for ideal wine quality. I think more future research need to be done to find its ideal quantity.
